\section{Discussion}
Looking at the requirements for the overall system, the final implementaion is a close attempt, however
minor functionality are still missing, specifically the suspend mechanism, which should halt the 
thread execution.

Looking at the Active object implementation, because the system is dealing with almost
instantenously executed code, there is no real consequence of execution time, which is
why an artificial thread sleep is introduced. This is also why the all the futures 
used in the implementation are void. They are merely used to indicate that some work has been done,
however they don't return anything. And all state processing is being held in the Servant.

As to whether the desired implementation lives up to the provided diagram is up for discussion,
some of the descriptors, specifically "configX", "eventX/responseM n eventX" seemed a 
little vague, so some artistic liberty was necessary to create a satisfactory solution.


\section{Conclusion}
It can then be concluded that the group has made a valid implementation of the requested system.
All critical functionality is preset, however some minor functions are 
still missing, such as the Suspend state in the Operational statemachine. As mentioned some artistic liberty has 
been taken due to vague assignment description, which in turn, has led to decisions like using 
void futures, as the provided state diagram was up to interpretation.


%skriv omkring besværligheder, og hvorfor suspend ikke er blevet implementeret.